---
title: "Coding_Challenge_5"
author: "Anet Filipova, Brynleigh Payne"
date: "2025-03-20"
output:
  md_document:
    variant: gfm
  html_document:
    toc: true
    toc_float: true
  pdf_document:
  word_document:
---
# Question 1: Read dataset
```{r}
# Clear memory
rm(list=ls(all = TRUE))

# Loading tidyverse library
library(tidyverse)

diversity <- read.csv("DiversityData.csv" , na.strings = "na")
meta <- read.csv("Metadata.csv", na.strings = "na")
```

# Question 2: Joining datasets by column Code
```{r}
alpha <- (left_join(meta, diversity, by = "Code")) # adding the richness data to the metadata based on on the common column of sampleID

```
# Question 3: Calculate Pielou’s evenness index: 
```{r}
# We are creating a new column that will be the result of shannon divided by log of richness
alpha_even <- mutate(alpha, logRich = shannon/log(richness))

```
# Question 4: Using the summarise() function to calculate the mean and standard error evenness grouped by crop over time.
```{r}
alpha_average <- alpha_even %>% # We start with this dataset
  select(Code, Crop, Time_Point, Replicate, Water_Imbibed, shannon, invsimpson, simpson, richness) %>%
  group_by(Crop, Time_Point) %>% # grouping by Crop and Tmepoint to later calculate summary stats by group
  mutate(logRich = shannon/log(richness)) %>% # creating a new column of the log richness
  summarise(Mean.even = mean(logRich), # calculating the mean evennes, stdeviation, and standard error
            n = n(), # output of the sample size per group
            sd.dev = sd(logRich)) %>% # calculating the standard deviation of logRich
  mutate(std.err = sd.dev/sqrt(n)) # calculating the standard error which will be stored in the column std.err
```

# Question 5: Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
```{r}
alpha_average2 <- alpha_average %>% # We start with this dataset
  select(Time_Point, Crop, Mean.even) %>%
  pivot_wider(names_from = Crop, values_from = Mean.even) %>% # pivot to wide format by taking the names from the Crop column and the values from the mean.even column
  mutate(diff.cotton.even = Soil - Cotton) %>% # calculate the difference between Soil and Cotton
  mutate(diff.soybean.even = Soil - Soybean)  # calculate the difference between Soil and Soybean
```

# Questio 6: Connecting it to plots
```{r}
alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even), names_to = "diff") %>%
  ggplot(aes(x = Time_Point, y = value, color = diff)) + # Plot it 
  geom_line() +
  theme_classic() +
  xlab("Time (hrs)") +
  ylab("Difference from soil in Pielou's evenness")
```

[Link to GitHub](https://github.com/AnetFilipova/Reproducible_Data)






